New and simplified manual controls for projection and slice tours, with application to exploring classification boundaries in high dimensions

### 2023-01-18 review/response.Rmd

---
title: "Response to review: New and simplified manual controls for projection and slice tours, with application to exploring classification boundaries in high dimensions"
author: "Ursula Laa, Alex Aumann, Dianne Cook, German Valencia"
date: "2023-01-05"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

 XXX If use new slice center control for fig 7, need to make a note of it here, updated the mathematica notebook and check text and captions correct.

@@

1- Your revision should not increase the length of the manuscript.

2- I will download and run your code as part of the next round of reviews.  I suggest, but don't require, the tidyverse style for your code and, in any case, ask that you submit code that is documented and as user-friendly as possible.  Before resubmitting please run your code in a "clean" environment as issues such as required packages that aren't loaded or calls to files that don't exist are common and can delay publication (I'm not saying these exist in your case, just asking you to be aware!).

3- Your paper will be printed in black and white.  The online version will be in color but to accommodate printing without color please be sure you don't refer to color in either the figure captions or text and that all figures are still interpretable in black and white.

\textcolor{blue}{Colour in figures has been changed to work in black and white as well as colour. Captions have been adjusted accordingly. TODO}

@@

- Reproducibility:
    - Code should be reproducible and not depend on user paths.
    - Data files should be uploaded along with code files (they don't seem to be part of supplemental material?)
    - Not having the ability to actually run the code (because the data wasn't included) makes it somewhat diﬀicult to assess the software side of this paper.
- Grammar: There are a lot of minor grammatical issues with this paper. I suggest utilizing a tool such as Grammarly to spot and fix them. The ones that I noticed as I read through the first section are detailed in minor issues below, but I am trying to focus more on the content than the style, so I will try to not nit-pick through the rest of the paper and trust that the authors can use the online tools.
- Cite your software packages appropriately - they're tools you depend on to do your work and should be cited in the same way papers are cited.

    – Page 3, line 28: Capitalize Mathematica and cite it appropriately.

    – Page 3, line 32: Cite R

    – Page 7, line 28: Cite Shiny

\textcolor{blue}{Done - the citations were there, but not placed at the first mention.}

- Notation

### 2023-01-18 paper/paper.Rmd
updated response

This paper explains the new manual controls for projection and slice tours. The next section describes the new algorithm for manual control, for both projections and slices. The use of these methods is illustrated to compare and contrast boundaries constructed by different classifiers. The software section describes a Mathematica @Mathematica package that is used for the application, and describes the interactive environment that would be desirable within R @rref as new technology becomes available. The paper is accompanied by an appendix with more details and adjustments to the manual controls, and three Mathematica notebooks that can be used to reproduce the application.

@@

The implementation of the manual tour as suggested here requires the visualization of the current projection in terms of an axis display (see Figure \ref{fig:manualsequence} for an example). This display should be able to track the mouse position and adjust the projection based on the user selection. The user can select one axis (corresponding to one variable) by clicking on it, and then adjust the position of that axis by dragging it during the click event. In practice the closest axis will be selected, and the dragging results in tracking of the mouse position which will update the projection in small steps, such that no further interpolation is required.

The interactions in the axis display need to be mapped back onto the current projection matrix, which will then be orthonormalized before feeding back into the axis display. A second, linked display shows the projected data in sync with the updates from the axis display.

In addition, we also want to be able to look at slices of the data and select slicing parameters interactively. Here we have implemented a switch to change between projection and slice view, a slider that adjusts the slice thickness $h$ (dubbed `height` in the `Mathematica` package) and a numeric input to specify the center point $c$ explicitly. The display will update based on these inputs, in particular the view will jump to a new slice when $c$ is changed (though an interpolation might also be useful).

One might implement such an interactive interface via R Shiny @shinym, but in particular the tracking of the mouse position in small increments might pose a challenge. Here we found the dynamic graphics interface from Mathematica to be especially useful. In the following we first describe the relevant Mathematica functionalities and how they were used in the implementation, followed by a brief sketch of how the new approaches are implemented in the `tourr` package. There is hope that better interactivity will be available for R soon, which might allow for implementation of the techniques developed in Mathematica.

@@

The R package tourr [@tourr] provides numerous types of tours. An additional function, `radial_tour()` has been constructed that will generate a sequence of projections that will decrease the coordinates for $V_m$ to $\boldsymbol{0}$ and back to the original values. It is applicable for any projection dimension, $d$. Changing slice center manually can be accomplished with the new function, `manual_slice()`. This changes the value of the center point ($\mathbf{c}$) of the slice, along a selected variable axis. It slices from the center of the data out to one side, back to the center, out to the opposite side, and then back to the center. The appendix contains a discussion of a possible visual guide for the slice center location.

>>>>

This paper explains the new manual controls for projection and slice tours. The next section describes the new algorithm for manual control, for both projections and slices. The use of these methods is illustrated to compare and contrast boundaries constructed by different classifiers. The software section describes a Mathematica [@Mathematica] package that is used for the application, and describes the interactive environment that would be desirable within R [@rref] as new technology becomes available. The paper is accompanied by an appendix with more details and adjustments to the manual controls, and three Mathematica notebooks that can be used to reproduce the applications.

@@

The implementation of the manual tour as suggested here requires the visualization of the current projection in terms of an axis display (see Figure \ref{fig:manualsequence} for an example). This display should be able to track the mouse position and adjust the projection based on the \textcolor{blue}{user's cursor position}. The user can select one axis (corresponding to one variable) by clicking on it, and then adjust the position of that axis by dragging it. In practice, the closest axis will be selected, and the dragging results in tracking of the mouse position which will update the projection in small steps, such that no further interpolation is required.

The interactions in the axis display need to be mapped back onto the current projection matrix, which will then be orthonormalized and \textcolor{blue}{re-displayed}. A second, linked display shows the projected data in sync with the updates from the axis display.

In addition, we also want to be able to look at slices of the data and select slicing parameters interactively. Here we have implemented a switch to change between projection and slice view, a slider that adjusts the slice thickness, $h$ (named `height` in the Mathematica interface), and a numeric input to specify the center point $c$ explicitly. The display will update based on these input values, in particular the view will jump to a new slice when $c$ is changed.<!-- (though an interpolation might also be useful).-->

\textcolor{blue}{One might implement such an interactive interface in R, for example using `shiny` [@shinym], with `plotly` [@plotly] drawing. It could be possible with `rgl` [@rgl] but this package is focused on 3D, not the arbitrary $p$-D of tours. The primary tool needed is the tracking of the cursor position in small increments and it is not readily available in R. The `plotly` package is designed to add on to `ggplot2` graphics, not to be a programmable interactive graphics system. It would have been possible with the `cranvas` system [@cranvas] or `rJava` [@rJava] but neither has a graphics toolkit that operates across platforms today. What we have chosen to do is to use the interactivity of Mathematica to experiment, and to implement as an animation in the `tourr` [@tourr] package. There is hope that better interactivity will be available for R soon, perhaps in javascript, which might allow for experimentation to develop new techniques as available in Mathematica.}

@@

The R package `tourr` provides numerous types of tours. An additional function, `radial_tour()` has been constructed that will generate a sequence of projections that will decrease the coordinates for $V_m$ to $\boldsymbol{0}$ and back to the original values. It is applicable for any projection dimension, $d$. Changing slice center manually can be accomplished with the new function, `manual_slice()`. This changes the value of the center point ($\mathbf{c}$) of the slice, along a selected variable axis. It slices from the center of the data out to one side, back to the center, out to the opposite side, and then back to the center. The appendix contains a discussion of a possible visual guide for the slice center location.

### 2023-01-18 paper/paper.tex
updated response

This paper explains the new manual controls for projection and slice
tours. The next section describes the new algorithm for manual control,
for both projections and slices. The use of these methods is illustrated
to compare and contrast boundaries constructed by different classifiers.
The software section describes a Mathematica \citet{Mathematica} package
that is used for the application, and describes the interactive
environment that would be desirable within R \citet{rref} as new
technology becomes available. The paper is accompanied by an appendix
with more details and adjustments to the manual controls, and three
Mathematica notebooks that can be used to reproduce the application.

@@

The implementation of the manual tour as suggested here requires the
visualization of the current projection in terms of an axis display (see
Figure \ref{fig:manualsequence} for an example). This display should be
able to track the mouse position and adjust the projection based on the
user selection. The user can select one axis (corresponding to one
variable) by clicking on it, and then adjust the position of that axis
by dragging it during the click event. In practice the closest axis will
be selected, and the dragging results in tracking of the mouse position
which will update the projection in small steps, such that no further
interpolation is required.

The interactions in the axis display need to be mapped back onto the
current projection matrix, which will then be orthonormalized before
feeding back into the axis display. A second, linked display shows the
projected data in sync with the updates from the axis display.

In addition, we also want to be able to look at slices of the data and
select slicing parameters interactively. Here we have implemented a
switch to change between projection and slice view, a slider that
adjusts the slice thickness \(h\) (dubbed \texttt{height} in the
\texttt{Mathematica} package) and a numeric input to specify the center
point \(c\) explicitly. The display will update based on these inputs,
in particular the view will jump to a new slice when \(c\) is changed
(though an interpolation might also be useful).

One might implement such an interactive interface via R Shiny
\citet{shinym}, but in particular the tracking of the mouse position in
small increments might pose a challenge. Here we found the dynamic
graphics interface from Mathematica to be especially useful. In the
following we first describe the relevant Mathematica functionalities and
how they were used in the implementation, followed by a brief sketch of
how the new approaches are implemented in the \texttt{tourr} package.
There is hope that better interactivity will be available for R soon,
which might allow for implementation of the techniques developed in
Mathematica.

@@

The R package tourr \citep{tourr} provides numerous types of tours. An
additional function, \texttt{radial\_tour()} has been constructed that
will generate a sequence of projections that will decrease the
coordinates for \(V_m\) to \(\boldsymbol{0}\) and back to the original
values. It is applicable for any projection dimension, \(d\). Changing
slice center manually can be accomplished with the new function,
\texttt{manual\_slice()}. This changes the value of the center point
(\(\mathbf{c}\)) of the slice, along a selected variable axis. It slices
from the center of the data out to one side, back to the center, out to
the opposite side, and then back to the center. The appendix contains a
discussion of a possible visual guide for the slice center location.

>>>>

This paper explains the new manual controls for projection and slice
tours. The next section describes the new algorithm for manual control,
for both projections and slices. The use of these methods is illustrated
to compare and contrast boundaries constructed by different classifiers.
The software section describes a Mathematica \citep{Mathematica} package
that is used for the application, and describes the interactive
environment that would be desirable within R \citep{rref} as new
technology becomes available. The paper is accompanied by an appendix
with more details and adjustments to the manual controls, and three
Mathematica notebooks that can be used to reproduce the applications.

@@

The implementation of the manual tour as suggested here requires the
visualization of the current projection in terms of an axis display (see
Figure \ref{fig:manualsequence} for an example). This display should be
able to track the mouse position and adjust the projection based on the
\textcolor{blue}{user's cursor position}. The user can select one axis
(corresponding to one variable) by clicking on it, and then adjust the
position of that axis by dragging it. In practice, the closest axis will
be selected, and the dragging results in tracking of the mouse position
which will update the projection in small steps, such that no further
interpolation is required.

The interactions in the axis display need to be mapped back onto the
current projection matrix, which will then be orthonormalized and
\textcolor{blue}{re-displayed}. A second, linked display shows the
projected data in sync with the updates from the axis display.

In addition, we also want to be able to look at slices of the data and
select slicing parameters interactively. Here we have implemented a
switch to change between projection and slice view, a slider that
adjusts the slice thickness, \(h\) (named \texttt{height} in the
Mathematica interface), and a numeric input to specify the center point
\(c\) explicitly. The display will update based on these input values,
in particular the view will jump to a new slice when \(c\) is changed.

\textcolor{blue}{One might implement such an interactive interface in R, for example using `shiny` [@shinym], with `plotly` [@plotly] drawing. It could be possible with `rgl` [@rgl] but this package is focused on 3D, not the arbitrary $p$-D of tours. The primary tool needed is the tracking of the cursor position in small increments and it is not readily available in R. The `plotly` package is designed to add on to `ggplot2` graphics, not to be a programmable interactive graphics system. It would have been possible with the `cranvas` system [@cranvas] or `rJava` [@rJava] but neither has a graphics toolkit that operates across platforms today. What we have chosen to do is to use the interactivity of Mathematica to experiment, and to implement as an animation in the `tourr` [@tourr] package. There is hope that better interactivity will be available for R soon, perhaps in javascript, which might allow for experimentation to develop new techniques as available in Mathematica.}

@@

The R package \texttt{tourr} provides numerous types of tours. An
additional function, \texttt{radial\_tour()} has been constructed that
will generate a sequence of projections that will decrease the
coordinates for \(V_m\) to \(\boldsymbol{0}\) and back to the original
values. It is applicable for any projection dimension, \(d\). Changing
slice center manually can be accomplished with the new function,
\texttt{manual\_slice()}. This changes the value of the center point
(\(\mathbf{c}\)) of the slice, along a selected variable axis. It slices
from the center of the data out to one side, back to the center, out to
the opposite side, and then back to the center. The appendix contains a
discussion of a possible visual guide for the slice center location.
